---
title: "DS710 Final Project -Ken Constable"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Load Libraries
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(ggformula)
library(readr)
library(tidyr)
library(lubridate)
library(plotly)
library(scales)
library(forecast)
```

# Import Tweet Data from Python
+ import tweet data from python
+ only keep fields needed for the analysis
```{r}
getTwitterData<-function(fpath){
  require(readr)

  df<-read_csv(fpath)%>%
      select(hashtag,date,retweet,clean_text,sentiment,polarity,followers_count)
  return(df)
}
```


# Plot histograms
+ create before/after release date histograms
```{r}
plotHistograms<-function(df,relase_date='2019-11-8'){
  require(dplyr)
  require(lubridate)
  require(scales)
  require(ggformula)


  #split df into before and after based on the release_date
  before<-filter(df,date<relase_date)%>%
           select(polarity)

  after<-filter(df,date>=relase_date)%>%
          select(polarity)

  #build the sub-title
  sub.title<-paste("Mean Polarity Before(",round(mean(before$polarity),2)
                  ,") and after(",round(mean(after$polarity),2) ,") release date")

  #create a before/after indicator for faceting the charts
  #remove neutral tweets
  df%>%
    mutate(period=ifelse(date>=relase_date,'(2) After','(1) Before'))%>%
    filter(polarity!=0)%>%
    gf_dhistogram(~polarity
                  ,color=~period
                  ,fill=~period
                  ,bins=9
                  ,binwidth = 0.25
                  ,title = "Sentiment Distribution"
                  ,subtitle = sub.title
                  ,xlab = "Polarity"
                  ,ylab = "Density")%>%
    gf_fitdistr(dist='dnorm',color=~period)%>%
    gf_facet_grid(period ~.)%>%
    gf_refine(scale_fill_manual(values=c("grey","#2E4500"),aesthetics = c("fill","color")))%>%
    gf_theme(theme_minimal())

}
```


# plot sentiment over time
+ plot tweet volumn by sentitment vs date
+ add 5 day rolling average polarity
+ add 5 day rolling averaage % positive tweets
```{r}
plotTweetSentiment<-function(df,date='2019-11-8',alpha=0.05){
  require(dplyr)
  require(plotly)
  require(lubridate)

  #get counts of tweets by sentiment,date
  #if there are no counts for a particular day, replace the NA with zero
  #calculate total tweet volume, % positive tweets,before and after indicator
  d1<-group_by(df,date,sentiment)%>%
    summarise(count=n())%>%
    spread(sentiment,count)%>%
    ungroup()%>%
    arrange(date)%>%
    mutate( negative       = ifelse(is.na(negative),0,negative)
           ,positive       = ifelse(is.na(positive),0,positive)
           ,neutral        = ifelse(is.na(neutral),0,neutral)
           ,volume         = negative+positive+neutral
           ,pos.percent    = positive/volume
           ,ma.pos.percent = ma(pos.percent,5)
           ,period         = ifelse(date>=ymd(date),"after","before"))


  #find max y height, and before/after dates for plotting the rectangles for before/after
   max_volume<-max(d1$volume)*1.10
   min_date<-min(d1$date)
   max_date<-max(d1$date)
   hashtag<-df$hashtag[1]


  #determine spacing for annotations - varies by hashtag
  if(hashtag=="DrSleep"){
    # release date annotation
     a <- list(
      x = ymd(date),
      y = max_volume*0.9,
      text = "Release Date",
      xref = "x",
      yref = "y",
      showarrow = TRUE,
      arrowhead = 5,
      ax=-75,
      ay=0
    )

     #coordinates of before/after text
     x_before=ymd(date)-4
     y_before=max_volume*0.50
     x_after=ymd(date)+10
     y_after=max_volume*0.50

  }

  if(hashtag=="KnivesOut"){
    # release date annotation
     a <- list(
      x = ymd(date),
      y = max_volume*0.98,
      text = "Release Date",
      xref = "x",
      yref = "y",
      showarrow = TRUE,
      arrowhead = 5,
      ax=-125,
      ay=0
    )

     #coordinates of before/after text
     x_before=ymd(date)-8
     y_before=max_volume*0.55
     x_after=ymd(date)+1
     y_after=max_volume*0.70

  }

  if(hashtag=="CharliesAngels"){
    # release date annotation
     a <- list(
      x = ymd(date),
      y = max_volume*0.98,
      text = "Release Date",
      xref = "x",
      yref = "y",
      showarrow = TRUE,
      arrowhead = 5,
      ax=-125,
      ay=0
    )

     #coordinates of before/after text
     x_before=ymd(date)-5
     y_before=max_volume*0.70
     x_after=ymd(date)+10
     y_after=max_volume*0.70

  }



  #before/after annotations
   before<-getSentimentProportions(df,type="before",date=date,alpha=alpha)
   after <-getSentimentProportions(df,type="after", date=date,alpha=alpha)


  #get the average polarity by date
  d2<-group_by(df,date)%>%
    summarize(avg_polarity=mean(polarity))%>%
    mutate(ma_avg_polarity=ma(avg_polarity,5))

  #join together, ungroup for plotting
  d3<-inner_join(d1,d2,by="date")%>%
    dplyr::ungroup()


  #create the plot
  plot_ly(d3,x=~date,y=~positive,type="bar",name="Positive",color =I('#466900'),opacity=0.75)%>%
    add_trace(d3,x=~date,y=~negative,type='bar',name="Negative", color = I('#7C4325'),opacity=0.75)%>%
    add_trace(d3,x=~date,y=~neutral,type="bar",name="Neutral",color = I('#A1C423'))%>%
    add_trace(d3,x=~date,y=~ma.pos.percent,type="scatter",mode="line",name="Avg % Positive (5day)",yaxis='y2',color = I('#2E4500'),opacity=0.75)%>%
  add_trace(d3,x=~date,y=~ma_avg_polarity,type="scatter",mode="line",name="Avg Polarity (5day)",yaxis='y2',color=I("orange"))%>%  #3FAAA1


  layout(title = paste('Tweet Sentiment & Volume:',hashtag),
          xaxis = list(title = "Date"),
          yaxis = list(side = 'left', title = 'Tweet Volume', showgrid = TRUE),barmode="stack",
          yaxis2= list(side = 'right', overlaying = "y", title = 'Polarity and % Positive', showgrid = FALSE),
          shapes= list(list(type="rect",x0=min_date, x1=ymd(date),y0=0,y1=max_volume,fillcolor="grey",opacity=0.05),
                  list(type='rect',x0=ymd(date),x1=ymd(max_date), y0=0,y1=max_volume,fillcolor="red", opacity=0.025)),
         annotations=a,
         legend=list(y=0))%>%
         add_annotations(x=x_before,y=y_before,text=before,xref="x",yref="y",showarrow=FALSE,ax=-20,ay=-20)%>%
         add_annotations(x=x_after,y=y_after,text=after,xref="x",yref="y",showarrow=FALSE,ax=-20,ay=-20)


}

```

# Sentiment scatterplot
+ scatterplot of polarity vs time and # of followers
+ included option to remove tweets by authors over max_followers lime- removes outliers for plotting
```{r}

plotTweetScatterPlot<-function(df,release.date,max_followers='all'){

  require(dplyr)
  require(scales)
  require(ggformula)
  require(lubridate)

  # remove accounts with followers > max_followers
  # necessary if there are a few really large followers (which really skew the size dimension in the chart)
  if(is.numeric(max_followers)){
    removed<-nrow(filter(df,followers_count>max_followers))
    df<-filter(df,followers_count<max_followers)
    caption_text =paste("Removed ",removed," tweet(s) from accounts with followers > ",comma_format()(max_followers))
  }else{
    caption_text = ""
  }


  #find max y height, and before/after dates for plotting the rectangles for before/after
   min_date<-min(df$date)
   max_date<-max(df$date)



  # create the chart
  # add a rectagle to to show the release date
  gf_point(df
         ,gformula = polarity~date
         ,color=~sentiment
         ,size=~followers_count/1000
         ,title = "Sentiment by Date"
         ,subtitle = "Size proportional to followers"
         ,caption = caption_text
         ,alpha = 0.5)%>%
  gf_refine(scale_color_manual(values=c("#7C4325","#A1C423","#466900")))%>%
  gf_refine(labs(size = "Followers \n (1,000)", color="Sentiment"))%>%
  gf_refine(annotate("rect",xmin=min_date,xmax=ymd(release.date),ymin=-1,ymax=1,alpha=0.05))%>%
  gf_refine(annotate("rect",xmin=ymd(release.date),xmax=max_date,ymin=-1,ymax=1,alpha=0.05,fill="red"))%>%
  gf_refine(annotate("text",x=ymd(release.date),y=1.15,label="Release\n Date",size=3,color="grey"))%>%
  gf_theme(theme_minimal())
}

```




# Test for proportions - by release date
```{r}
testProportionsByReleaseDate<-function(df,release_date,alpha=0.05){
  #http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/R/R6_CategoricalDataAnalysis/R6_CategoricalDataAnalysis6.html


  #H0: p(positive,negative,neutral) before release = p(postive,negative,neutral) after release
  #Ha: p(before) !=p(after) - for at least one category

  #split the dataset by release_date
  df_before<-filter(df,date<ymd(release_date))%>%
    group_by(sentiment)%>%
    summarize(count=n())%>%
    arrange(sentiment)


  df_after<-filter(df,date>=ymd(release_date))%>%
    group_by(sentiment)%>%
    summarize(count=n())%>%
    arrange(sentiment)

  # some error handling, usually when you put the wrong date in for a movie
  # perform the test
  if(length(df_before)!=length(df_after)){
    print("Check the dates! Before length not equal to after length")
  }else{
    result<-prop.test(df_before$count,df_after$count,alternative = "two.sided")
    if(result$p.value<alpha){
      #reject Ho
      decision<-paste("Reject the Null, Proportions are different before/after relase date at an alpha level of ",alpha)
      #return(result)
    }else{
      #fail to reject ho
      decision<-paste("Fail to reject the null, proportions may be the same before/after release date at an alpha level of ",alpha)
      #return(result)
    }

  }

  #Return the test result, the before/after proportions, and the decision in the list
  return(list(result=result,before=df_before,after=df_after,decision=decision))

}



```


# Splits a dataframe by the date
+ used in the plotTweetSentiment() chart
+ converts proportions to formated text for plotting
```{r}
getSentimentProportions<-function(df,type,date="2019-11-8",alpha=0.05){

  #use the testProportionsByReleaseDate function to get the split of before/after proportions
  tmp<-testProportionsByReleaseDate(df,date,alpha)

  #convert the proportions to a formatted string for plotting
  if(type=="before"){
    df<-data.frame(sentitment=tmp[['before']][1],count=tmp[['before']][2])
    result<-'</br><b>Before</b></br>'
    total<-sum(df['count'])

    for(row in 1:nrow(df)){
      sent<-df[row,"sentiment"]
      per<-df[row,"count"]/total
      result<-paste(result,sent,":",scales::percent_format()(per),sep=' ','</br> ')
    }
  }else{
    df<-data.frame(sentitment=tmp[['after']][1],count=tmp[['after']][2])
    result<-'</br><b>After</b></br>'
    total<-sum(df['count'])

    for(row in 1:nrow(df)){
      sent<-df[row,"sentiment"]
      per<-df[row,"count"]/total
      result<-paste(result,sent,":",scales::percent_format()(per),sep=' ','</br> ')
    }
  }

  #return the formatted string
  return(result)
}


```



# Test for Proportions - by followers
+ this splits the tweets by the median (or input) follower count & performs the test of proportions by followers
```{r}
testProportionsByFollowers<-function(df,alpha=0.05,followers="median"){


  #H0: p(positive,negative,neutral) for non-influencers = p(postive,negative,neutral) for influencers
  #Ha: p(non-influencers) !=p(influencers) - for at least one category

  # determine how to split followers.Default calculates the median
  # else, use the limit provided
  if(followers=="median"){
    med.followers<-median(df$followers_count)
  }else{
    med.followers<-followers
  }

  #split the dataset by followers
  non_influencers<-filter(df,followers_count<med.followers)%>%
    group_by(sentiment)%>%
    summarize(count=n())%>%
    arrange(sentiment)


  influencers<-filter(df,date>=med.followers)%>%
    group_by(sentiment)%>%
    summarize(count=n())%>%
    arrange(sentiment)


  # perform the test.
  if(length(non_influencers)!=length(influencers)){
    print("Sentiment missing from influencer or non-influencers proportions")
  }else{
    result<-prop.test(non_influencers$count,influencers$count,alternative = "two.sided")
    if(result$p.value<alpha){
      #reject Ho
      decision<-paste("Reject the Null, Proportions are different for influencers at an alpha level of ",alpha)
      #return(result)
    }else{
      #fail to reject ho
      decision<-paste("Fail to reject the null, proportions may be the same for influencers/non-influencers at an alpha level of ",alpha)
      #return(result)
    }

  }

  return(list(result=result,non_influencers=non_influencers,influencers=influencers,decision=decision))

}
```

# T-Test for means
+ tests of two means are the same
+ used to test mean polarity before/after the movie release
```{r}
testForMeans<-function(df,date,alpha=0.05){

  #split the dataset by date
  before<-mutate(df,period=ifelse(date<ymd("2019-11-8"),"before","after"))%>%
    filter(period=="after")

  after<-mutate(df,period=ifelse(date<ymd("2019-11-8"),"before","after"))%>%
    filter(period=="after")

  #perform the test of means
  result<-t.test(before$polarity,after$polarity,alternative = "two.sided")

  if(result$p.value>alpha){
    #fail to reject the null
    print("Fail to reject the null=> There may be no differnce in means")

  }else{
    #reject the nulll
    print("Reject the Null=> There is a difference in means")
  }

  return(result)

}
```




# Get data from python
+import the tweet data from python for the two hashtags
```{r message=FALSE, warning=FALSE}
drSleep<-getTwitterData("TweetData/Output_DrSleep.csv")
knivesOut<-getTwitterData("TweetData/Output_KnivesOut.csv")

```

# plot the results for DrSleep
```{r message=FALSE, warning=FALSE}
plotTweetSentiment(drSleep,"2019-11-8")
plotTweetScatterPlot(drSleep,max_followers = 13000000,release.date =  "2019-11-8")
plotHistograms(drSleep,"2019-11-8")
testProportionsByReleaseDate(drSleep,"2019-11-8")  #significant - proportions are different
testProportionsByFollowers(drSleep)   #fail to reject the null, no difference in proportions for influencers/non-influencers
testForMeans(drSleep,"2019-11-8")     #fail to reject the null, no differenc in mean polarity before/after release date

```

# Plot the results for KnivesOut
```{r message=FALSE, warning=FALSE}
plotTweetSentiment(knivesOut,"2019-11-27")
plotTweetScatterPlot(knivesOut,max_followers = "all",release.date =  "2019-11-27")
plotHistograms(knivesOut,"2019-11-27")
testProportionsByReleaseDate(knivesOut,"2019-11-27")  #sigficiant -there is a differece in proportions before/after release date
testProportionsByFollowers(knivesOut)    #significant -there is a difference in proportions for influencers/non-influences
testForMeans(knivesOut,"2019-11-26")     #fail to reject the null -no difference in mean polarity before/after release date

```
